---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
First, I read in the dataset.

```{r,echo=TRUE,results='hide'}
dat <- read.csv("activity.csv")
library(dplyr)
library(lubridate)
dat$date <- ymd(dat$date)
```

## What is mean total number of steps taken per day?
Next, I find the mean and median number of steps taken in a day.

```{r,echo=TRUE}
totalstepsbyday <- tapply(dat$steps,dat$date,sum,na.rm=TRUE)
hist(totalstepsbyday)
meanstepsperday <- mean(totalstepsbyday)
medianstepsperday <- median(totalstepsbyday)
```

The mean of steps taken per day is `r meanstepsperday` 
and the median of steps taken per day is `r medianstepsperday`.

## What is the average daily activity pattern?
Let's look at the daily activity pattern.

```{r, echo=TRUE}
averagestepsperinterval <- tapply(dat$steps,dat$interval,mean,na.rm=TRUE)
plot(dat$interval[1:288], averagestepsperinterval, type = 'l', xlab = "Interval", ylab="Average Steps")
maxinterval <- dat$interval[which.max(averagestepsperinterval)]
```

The 5-minute interval with the most average steps is the one starting at `r maxinterval`.

## Imputing missing values
Let's work with some of our missing values.

```{r,echo=TRUE}
numberofmissingvalues <- sum(is.na(dat$steps))
```



There are `r numberofmissingvalues` observations with missing values.
Let's replace those with the average number of steps for any given interval.

```{r, echo=TRUE}
datnew <- dat
datnew$steps[is.na(datnew$steps)] <- mean(datnew$steps, na.rm = T)
colSums(is.na(datnew))
totalstepsbydaynew <- tapply(datnew$steps,datnew$date,sum,na.rm=TRUE)
hist(totalstepsbydaynew)
meanstepsperday2 <- mean(totalstepsbyday)
medianstepsperday2 <- median(totalstepsbyday)
```

The mean of steps taken per day is `r meanstepsperday2` 
and the median of steps taken per day is `r medianstepsperday2`.

The distribution became less skewed with the imputed missing values.


## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
datnew$week <- weekdays(datnew$date)
datnew$week  <- ifelse(datnew$week %in% c("Saturday","Sunday"), "Weekend","Weekday")
datnew$week <- as.factor(datnew$week)
library(lattice)
averagestepsperinterval2 <- tapply(datnew$steps,list(datnew$interval,datnew$week),mean,na.rm=TRUE)
par(mfrow=c(1,2))
plot(dat$interval[1:288], averagestepsperinterval2[,1], type = 'l', xlab = "Interval", ylab="Average Steps", main ="Weekday")
plot(dat$interval[1:288], averagestepsperinterval2[,2], type = 'l', xlab = "Interval", ylab="Average Steps", main="Weekend")
```



